04_analysis

### 2023-10-16 04_analysis.Rmd
Update 04_analysis.Rmd

Required Packages
```{r}
library(tidyverse)
library(mgcv)
```

Read in RDS files of each buoy + all environ variables + AIS (pending)
```{r}
ftlist <- dir("data/", recursive=TRUE, full.names=TRUE, pattern="fulltrack")

for (ft in ftlist) {
  df <- readRDS(ft)
  assign(paste0("fulltrack", substr(ft, 15, 16)), df)
}

@@

Bind ALL DRIFTS into one dataframe
```{r}
allDrifts <- rbind(fulltrack07, fulltrack08, fulltrack10, fulltrack12, fulltrack13, fulltrack14, 
                   fulltrack16, fulltrack18, fulltrack19, fulltrack20, fulltrack21, fulltrack22, fulltrack23)
```

### Third octave levels histograms

```{r all-drifts-all-freq-histogram}
# create list of fulltrack dfs from global environment to loop through
dlist <- grep(pattern = "fulltrack", names(.GlobalEnv), value = T)

# all Drifts - all TOL bands, facet wrap
for (ft in dlist) {
pivot1 <- get(ft) %>%
  pivot_longer(cols = starts_with("TOL") | starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

print(ggplot(data = pivot1, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of third octave level intensities: Drift ", substr(ft,10,11))) +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  (facet_grid(~factor(SL, levels = c('BB_20.24000', 'TOL_63', 'TOL_125', 'TOL_2000', 'TOL_5000', 'TOL_20000')))))
}

# Single plot for ALL DRIFTS
pivot3 <- allDrifts %>%
  pivot_longer(cols = starts_with("TOL") | starts_with("BB"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

#SL$size_f <- factor(SL$ , levels = c('BB_20.2400', 'TOL_63', 'TOL_125', 'TOL_2000', 'TOL_5000', 'TOL_20000'))))

ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 3, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: All Drifts Combined") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~factor(SL, levels = c('BB_20.24000', 'TOL_63', 'TOL_125', 'TOL_2000', 'TOL_5000', 'TOL_20000')))

## ignore??
```{r drift-8-histogram}
#ggplot(data = pivot3, aes(Hz, fill = factor(BWpresence))) +
#  geom_histogram(binwidth = 1, color = "black") +
#  theme_classic() +
#  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: Buoy 8") +
#  scale_x_continuous(name = "Sound Intensity (dB)") +
#  scale_fill_manual(name = "Beaked Whale Presence", 
#                    labels = c("0" = "Absent", "1" = "Present"),
#                    values = c("0" = "bisque1",
#                               "1" = "brown4")) +
#  theme(legend.position = c(0.6,0.7)) +
#  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

```{r all-drifts-two-freq-histogram}
# All Drifts - TOL bands 125 and 2000 only

TOL.labs <- c(`TOL_125` = "TOL at 125 Hz", 
              `TOL_2000` = "TOL at 2000 Hz") #creates label names for plots below

for (ft in dlist) {
pivot2 <- get(ft) %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

print(ggplot(data = pivot2, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of Third-Octave Level (TOL) Intensities: Drift ", substr(ft,10,11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs)))
}

# Single plot for ALL DRIFTS combined
pivot4 <- allDrifts %>%
  dplyr::select(-TOL_63, -TOL_20000, -TOL_5000) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "TOL", values_to = "Hz")    # Pivot longer for histogram facet grid

ggplot(data = pivot4, aes(Hz, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle("Distribution of Third-Octave Level (TOL) Intensities: All Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_manual(name = "Whale Presence", 
                    labels = c("0" = "Absent", "1" = "Present"),
                    values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.6,0.7)) +
  facet_wrap(~TOL, labeller = as_labeller(TOL.labs))
```

Timeline of drifts in ocean
```{r}
ggplot(data = allDrifts, aes(x = UTC, y = station*(-1))) + geom_point(aes(color = factor(station)))

#Note the break in drift 20 - I think it wasn't transmitting for several days (11/15-11/19)
```

Avg sound levels by band (scatterplot)
```{r}
avgdB <- TO_23 %>%
  summarize(TOL_25 = mean(TOL_25),
            TOL_31.5 = mean(TOL_31.5),
            TOL_40 = mean(TOL_40),
            TOL_50 = mean(TOL_50),
            TOL_63 = mean(TOL_63),
            TOL_80 = mean(TOL_80),
            TOL_100 = mean(TOL_100),
            TOL_125 = mean(TOL_125),
            TOL_160 = mean(TOL_160),
            TOL_200 = mean(TOL_200),
            TOL_250 = mean(TOL_250),
            TOL_315 = mean(TOL_315),
            TOL_400 = mean(TOL_400),
            TOL_500 = mean(TOL_500),
            TOL_630 = mean(TOL_630),
            TOL_800 = mean(TOL_800),
            TOL_1000 = mean(TOL_1000),
            TOL_1250 = mean(TOL_1250),
            TOL_1600 = mean(TOL_1600),
            TOL_2000 = mean(TOL_2000),
            TOL_2500 = mean(TOL_2500),
            TOL_3150 = mean(TOL_3150),
            TOL_4000 = mean(TOL_4000),
            TOL_5000 = mean(TOL_5000),
            TOL_6300 = mean(TOL_6300),
            TOL_8000 = mean(TOL_8000),
            TOL_10000 = mean(TOL_10000),
            TOL_12500 = mean(TOL_12500),
            TOL_16000 = mean(TOL_16000),
            TOL_20000 = mean(TOL_20000)) %>%
  pivot_longer(cols = TOL_25:TOL_20000,
               names_to = "frequency",
               values_to="dB")

ggplot(avgdB, aes(x = factor(frequency, levels = unique(frequency)), y = dB)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Box plots
```{r}
TO_longer <- pivot_longer(TO_10, cols = TOL_25:TOL_20000,
               names_to = "frequency",
               values_to="dB")
ggplot(TO_longer, aes(x = factor(frequency, levels = unique(frequency)), y = dB)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Violin plot
ggplot(TO_longer, aes(x = factor(frequency, levels = unique(frequency)), y = dB)) + 
  geom_violin() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

@@

# SST over time (strange plot)
ggplot(allDrifts, aes(x = UTC, y = sst_mean)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Depth vs TOL 125
ggplot(data = allDrifts, aes(x = depth, y = `TOL_125`)) +
  geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

#Mean depth over time
#ggplot(means_08, aes(x = `as.Date(UTC)`, y = depth*(-1))) +
#  geom_line()

@@

Facet Grids for Env. Variables
```{r}
#Get avg daily environmental variables
means <- allDrifts %>%
  group_by(as.Date(UTC))%>%
  summarize(sst_mean = mean(sst_mean),
            dist_slope = mean(dist2slope),
            depth = mean(depth),
            slope = mean(bathy_slope),
            dist_shore = mean(dist2shore),
            chlorophyll_mean = mean(chlorophyll_mean),
            ssh_mean = mean(ssh_mean),
            mldDepth = mean(mldDepth),
            mldTemp = mean(mldTemp),
            ttDepth = mean(ttDepth),
            ttTemp = mean(ttTemp),
            temp400 = mean(temp400),
            sal400 = mean(sal400),
            curlmean = mean(curl_mean),
            Wh_presence = sum(BWpresence))
```

@@

NEEDS FIXING/LOOP/AIS DATA FOR OTHER DRIFTS
Co-occurrence plots of drift 8 and vessels 
```{r}
#### 09/04/18 co-occurrence of vessel and drift within 10 km
track08_09.04 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-04 00:00:00" & UTC <= "2018-09-05 00:00:00")

ggplot(data = track08_09.04, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-04 14:13:00"), color = "yellow", linewidth = 2) +
  geom_text(aes(as.POSIXct("2018-09-04 14:13:00"), 88, label = "Enter vessel", hjust = 1.0)) +
  geom_vline(xintercept = as.POSIXct("2018-09-04 15:17:00"), color = "yellow", linewidth = 2) +
  geom_text(aes(as.POSIXct("2018-09-04 15:13:00"), 87, label = "Exit vessel", hjust = -0.1)) +
  geom_line() + 
  geom_point(data = track08_09.04, aes(color = species, size = 2)) +
  scale_color_discrete(na.translate=FALSE) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-04-18\n(Within 7 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/05/18 co-occurrence of vessel and drift within 10 km.
track08_09.05 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-04 21:00:00" & UTC <= "2018-09-06 00:00:00")

ggplot(data = track08_09.05, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-05 01:33:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-05 01:33:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-05 02:08:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-05 02:08:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() + 
  geom_point(data = track08_09.05, aes(color = species, size = 2)) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-05-18\n(Within 5 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/10/18 co-occurrence of vessel and drift within 10 km.
track08_09.10 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-10 00:00:00" & UTC <= "2018-09-11 00:00:00")

ggplot(data = track08_09.10, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-10 12:58:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-10 12:58:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-10 14:01:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-10 14:01:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() +
  geom_point(data = track08_09.10, aes(color = species, size = 2)) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-10-18\n(Within 2 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/22/18 co-occurrence of vessel and drift within 10 km
track08_09.22 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-22 00:00:00" & UTC <= "2018-09-23 00:00:00")

ggplot(data = track08_09.22, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-22 19:13:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-22 19:13:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-22 19:34:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-22 19:34:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() + 
  geom_point(data = track08_09.22, aes(color = species, size = 2)) +
  scale_size_continuous(guide = "none") +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-22-18\n(Within 9 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

#### 09/23/18 co-occurrence of vessel and drift within 10 km. NO whale detections in this time period.
track08_09.23 <- tracks_SS_08 %>%
  filter(UTC >= "2018-09-23 00:00:00" & UTC <= "2018-09-24 00:00:00")

ggplot(data = track08_09.23, aes(x = UTC, y = `TOL_125`)) +
  geom_vline(xintercept = as.POSIXct("2018-09-23 03:55:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-23 03:55:00"), 86, label = "Enter vessel", hjust = 1)) +
  geom_vline(xintercept = as.POSIXct("2018-09-23 04:23:00"), color = "yellow", linewidth = 1) +
  geom_text(aes(as.POSIXct("2018-09-23 04:23:00"), 85, label = "Exit vessel", hjust = -0.1)) +
  geom_line() +
  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-23-18\n(Within 6 km)") +
  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

IS THIS HELPFUL??
Whale Occurrences
```{r}
# Filter Drift 8 by Whale presence and absence
W1 <- tracks_SS_08 %>%
  filter(Wpresence == 1)
W0 <- tracks_SS_08 %>%
  filter(Wpresence == 0)

ggplot(data = tracks_SS_08, mapping = aes(TOL_125)) + geom_histogram(binwidth = 0.1, color = "black")

ggplot() + 
  geom_histogram(data = W0, aes(TOL_5000), binwidth = 0.5, fill = "black", color = "white") +
  geom_histogram(data = W1, aes(TOL_5000), binwidth = 0.5, fill = "yellow", color = "black")

# Percent of time BWs are detected
print(paste(nrow(W1) / nrow(tracks_SS_08) * 100,"%"))

# Most abundant whale species when whale is detected (remove NA values)
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot() + geom_bar(aes(species))

# Sound levels whales are exposed to by species
ZC <- allDrifts %>%
  filter(species == "ZC")

BW43 <- allDrifts %>%
  filter(species == "BW43")

BB <- allDrifts %>%
  filter(species == "BB")

BWC <- allDrifts %>%
  filter(species == "BWC")

BW37V <- allDrifts %>%
  filter(species == "BW37V")

PM <- allDrifts %>%
  filter(species == "PM")

print(max(ZC$TOL_125))
print(max(BW43$TOL_125))
print(max(BB$TOL_125))
print(max(BWC$TOL_125))
print(max(BW37V$TOL_125))
print(max(PM$TOL_125))

ggplot() + geom_histogram(data = ZC, mapping = aes(TOL_125), binwidth = 0.2, fill = "brown", color = "white") +
  geom_histogram(data = BW43, mapping = aes(TOL_125), binwidth = 0.2, fill = "yellow", color = "white") +
  theme_classic()
```

Density Plots - Species at TOL sound levels
```{r density-plots}
# Broadband
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(`BB_20.24000`, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("Broadband Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# Broadband SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(`BB_20.24000`, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("Broadband ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 63
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_63, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 63, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 63 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_63, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 63 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 125
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_125, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 125, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

@@

  print(ggplot(data = plt, aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 2000, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 2000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 2000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 5000
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_5000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 5000, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 5000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_5000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 5000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))

# TOL 20,000
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_20000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 20000, Drift ", substr(ft, 10, 11))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 20,000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_20000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 20,000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

ARE THESE HELPFUL??
Histograms with whale presence/absence and environmental variables. Doesnt differentiate by species though
```{r}
# SST
ggplot(data = allDrifts, aes(sst_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.1, color = "black")

# Depth
ggplot(data = allDrifts, aes(depth, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 100, color = "black")

# Bottom slope
ggplot(data = allDrifts, aes(bathy_slope, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 1, color = "black")

# SSH
ggplot(data = allDrifts, aes(ssh_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.005, color = "black")

# Chlorophyll
ggplot(data = allDrifts, aes(chlorophyll_mean, fill = factor(BWpresence))) +
  geom_histogram(binwidth = 0.05, color = "black")

>>>>

Required Packages
```{r setup}
library(tidyverse)
library(mgcv)
```

Read in RDS files of each buoy + all environ variables + AIS
```{r}
ftlist <- dir("D:/Buoy_dfs/", recursive=TRUE, full.names=TRUE, pattern="GPSwhaleAISenvSS")

for (ft in ftlist) {
  df <- readRDS(ft)
  assign(paste0("GPSwhaleAISenvSS", substr(ft, 29, 30)), df)
}
```

Bind ALL DRIFTS into one dataframe
```{r}
dlist <- grep(pattern = "GPSwhaleAISenvSS", names(.GlobalEnv), value = T)

# Create a new column that specifies the drift number before the bind.
for (col in dlist){
  data <- get(col) %>% mutate(Deployment = substr(col,17,18))
  assign(paste0("GPSwhaleAISenvSS", substr(col,17,18)), data)
}

# Bind and save allDrifts dataframe
allDrifts <- rbind(GPSwhaleAISenvSS07, GPSwhaleAISenvSS08, GPSwhaleAISenvSS10, GPSwhaleAISenvSS12,
                   GPSwhaleAISenvSS13, GPSwhaleAISenvSS14, GPSwhaleAISenvSS16, GPSwhaleAISenvSS18,
                   GPSwhaleAISenvSS19, GPSwhaleAISenvSS20, GPSwhaleAISenvSS21, GPSwhaleAISenvSS22,
                   GPSwhaleAISenvSS23)
save(allDrifts, file = "data/allDrifts.rda")
```

### Third octave levels histograms
```{r each-drift-histogram}
# create list of fulltrack dfs from global environment to loop through
dlist <- grep(pattern = "GPSwhaleAISenvSS", names(.GlobalEnv), value = T)

# all Drifts - all TOL bands, facet wrap
for (ft in dlist) {
pivot1 <- get(ft) %>%
  pivot_longer(cols = starts_with("TOL"), names_to = "SL", values_to = "Hz")    # Pivot longer for histogram facet grid

print(ggplot(data = pivot1, aes(Hz, fill = factor(Wpresence))) +
  geom_histogram(binwidth = 1, color = "black") +
  theme_classic() +
  ggtitle(paste0("Distribution of TOL intensities: Drift", substr(ft,17,18))) +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.5,0.8)) +
  (facet_grid(~factor(SL, levels = c('TOL_125', 'TOL_2000')))))
}
```

```{r all-drifts-histogram}
# Single plot for ALL DRIFTS
pivot3 <- allDrifts %>%
  # Pivot longer for histogram facet grid
  pivot_longer(cols = starts_with("TOL"), names_to = "SL", values_to = "Hz")

ggplot(data = pivot3, aes(Hz, fill = factor(Wpresence))) +
  geom_histogram(binwidth = 2, color = "black") +
  theme_classic() +
  ggtitle("Distribution of third octave level intensities: All Drifts Combined") +
  labs(x = "Sound Intensity (dB)") +
  guides(fill = guide_legend(title = "Whale Presence")) +
  scale_fill_manual(values = c("0" = "bisque1",
                               "1" = "brown4")) +
  theme(legend.position = c(0.7,0.8)) +
  facet_grid(~factor(SL, levels = c('TOL_125', 'TOL_2000')))
```

Timeline of drifts in ocean
```{r}
ggplot(data = allDrifts, aes(x = UTC, y = Deployment)) + geom_point(aes(color = factor(Deployment)))
```

Looking at TOl2000 sound levels for each buoy and where Whale Detections happen
```{r fig.width=10}
hourly <- GPSwhaleAISenvSS10 %>% group_by(time = floor_date(UTC, '4 hours')) %>%
  summarize(mean_TOL2000 = mean(TOL_2000),
            sum_AIS = sum(shipin10km),
            sum_Wh = sum(Wpresence))

ZC_filt <- GPSwhaleAISenvSS10 %>% 
  filter(species %in% c("ZC", NA)) %>%
  mutate(rollAvg = zoo::rollapply(Wpresence, 6, mean, fill = NA))

ggplot(hourly, aes(x = time, y = mean_TOL2000)) + geom_line() + 
  geom_smooth() + 
  geom_point() +
  #geom_point(aes(color = species)) +
  #geom_vline(xintercept = hourly$UTC[hourly$shipin10km == 1], color = "yellow", alpha = 0.5) +
  scale_color_discrete(na.translate=FALSE) #+
 # scale_x_datetime(date_breaks = "1 day",limits = as.POSIXct(c("2018-09-24 00:00:00", "2018-10-03 00:00:00")))

@@

Compare overall TOL2000sound levels when whales are absent vs. present
```{r}
ggplot(allDrifts, aes(factor(Wpresence), TOL_2000)) + geom_boxplot()
```

@@

# SST over time (strange plot)
ggplot(allDrifts, aes(x = UTC, y = sst_mean)) +
  #geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)
  #geom_smooth(method = "loess")

# Depth vs TOL 125
ggplot(data = allDrifts, aes(x = `TOL_125`, y = depth *(-1))) +
  #geom_line() +
  geom_point(aes(color = species)) +
  scale_color_discrete(na.translate=FALSE)

@@

Facet Grids for Env. Variables
```{r}
#Get avg daily environmental variables
means <- allDrifts %>%
  group_by(as.Date(UTC))%>%
  summarize(sst_mean = mean(sst_mean),
            dist_slope = mean(dist2slope),
            depth = mean(depth),
            slope = mean(bathy_slope),
            dist_shore = mean(dist2shore),
            chlorophyll_mean = mean(chlorophyll_mean),
            ssh_mean = mean(ssh_mean),
            mldDepth = mean(mldDepth),
            mldTemp = mean(mldTemp),
            ttDepth = mean(ttDepth),
            ttTemp = mean(ttTemp),
            temp400 = mean(temp400),
            sal400 = mean(sal400),
            curlmean = mean(curl_mean),
            Wh_presence = sum(Wpresence))
View(means)
```

@@

Use as reference for similar plot (potentially??)
Co-occurrence plots of drift 8 and vessels 
```{r}
#### 09/04/18 co-occurrence of vessel and drift within 10 km
#track08_09.04 <- tracks_SS_08 %>%
#  filter(UTC >= "2018-09-04 00:00:00" & UTC <= "2018-09-05 00:00:00")

#ggplot(data = track08_09.04, aes(x = UTC, y = `TOL_125`)) +
#  geom_vline(xintercept = as.POSIXct("2018-09-04 14:13:00"), color = "yellow", linewidth = 2) +
#  geom_text(aes(as.POSIXct("2018-09-04 14:13:00"), 88, label = "Enter vessel", hjust = 1.0)) +
#  geom_vline(xintercept = as.POSIXct("2018-09-04 15:17:00"), color = "yellow", linewidth = 2) +
#  geom_text(aes(as.POSIXct("2018-09-04 15:13:00"), 87, label = "Exit vessel", hjust = -0.1)) +
#  geom_line() + 
#  geom_point(data = track08_09.04, aes(color = species, size = 2)) +
#  scale_color_discrete(na.translate=FALSE) +
#  scale_size_continuous(guide = "none") +
#  ggtitle("Sound levels during co-occurrence of Drift 8 and a vessel on 09-04-18\n(Within 7 km)") +
#  scale_y_continuous(name = "Sound Intensity (dB) at TOL 125 Hz") +
#  scale_color_discrete(na.translate=FALSE, labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

Density Plots - Species at TOL sound levels
```{r density-plots}
# TOL 125
for (ft in dlist) {
plt <- get(ft) %>%
  filter(!is.na(species))

  print(ggplot(data = plt, aes(TOL_125, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 125, Drift ", substr(ft, 17, 18))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

@@

  print(ggplot(data = plt, aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle(paste0("TOL 2000, Drift ", substr(ft, 17, 18))) +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43")))
}

# TOL 2000 SINGLE PLOT ALL DRIFTS
allDrifts %>%
  filter(!is.na(species)) %>%
  ggplot(aes(TOL_2000, fill = species)) + 
  geom_density(alpha = 0.5) +
  ggtitle("TOL 2000 ALL Drifts Combined") +
  scale_x_continuous(name = "Sound Intensity (dB)") +
  scale_fill_discrete(labels = c("ZC" = "Cuvier's", "BW43" = "BW43"))
```

Visualize SS for buoy 20 when GPS is malfunctioning
```{r}
# Visualize where detections are missing
GPSwhaleAISenvSS20$missingGPS <-ifelse(is.na(GPSwhaleAISenvSS20$Latitude),
                                       "yes", "no")
# Need to create rollnoise column
#ggplot(GPSwhaleAISenvSS20, aes(UTC, TOL_125)) + 
#  geom_point(aes(color = missingGPS)) + geom_line(aes(y=rollnoise))

Does time elapsed affect whale presence?

Linear model
```{r}
model6 <- lm(Wpresence>0 ~ timeAfterShip, data = allDrifts)
summary(model6)
plot(model6)
```

Avg. number of hours after ship encounter?
```{r}
mean(allDrifts$timeAfterShip) / 60 / 60
```

## Soundscape before, during, after ship passage?
Elevated TOL_125 sound levels with ship presence, only slight difference with TOL_2000
```{r}
# During
ship_event <- allDrifts %>% filter(shipin10km == 1)
no_ship <- allDrifts %>% filter(shipin10km == 0)

# Plot TOL 125
ggplot(allDrifts, aes(factor(shipin10km), TOL_125)) + geom_boxplot() + labs(x = "Ship Present?")
summary(ship_event$TOL_125)
summary(no_ship$TOL_125)

# Plot TOL_2000
ggplot(allDrifts, aes(factor(shipin10km), TOL_2000)) + geom_boxplot() + labs(x = "Ship Present?")
summary(ship_event$TOL_2000)
summary(no_ship$TOL_2000)

## How to show BEOFRE AND AFTER???
# create factors by days. discrete values like 10,000 - 20,000 seconds.
```

How often are there detections immediately following ship passage?
```{r}
# Filter by whale detections and mutate to hrs + days after ship event
detections <- allDrifts %>% 
  filter(Wpresence == 1) %>%
  mutate(hrsAfterShip = timeAfterShip/3600) %>%
  mutate(daysAfterShip = hrsAfterShip/24)

# Plot number of days after ship event by species
ggplot(detections, aes(daysAfterShip, species)) + geom_boxplot()
summary(detections$daysAfterShip) # Avg 5.4 days after ship
```
Are whales present before a ship passage?
```{r}
detect_b4_ship <- allDrifts %>%
  filter(Wpresence == 1) %>%
  mutate(hrsBeforeShip = timeBeforeShip/3600) %>%
  mutate(daysBeforeShip = hrsBeforeShip/24)

ggplot(detect_b4_ship, aes(daysBeforeShip)) + geom_histogram()
summary(detect_b4_ship$daysBeforeShip) # Avg 5.9 days before ship
```
Visualize whale detections and ship events
```{r fig.width=10}
detsOnly <- GPSwhaleAISenvSS08 %>% filter(Wpresence == 1)
shipOnly <- GPSwhaleAISenvSS08 %>% filter(shipin10km == 1)
